---
title: "Untitled"
author: "Leah Crowe"
date: '2023-06-05'
output: pdf_document
---

```{r setup, include=FALSE}
library(dplyr)
library(lubridate)
library(stringr)
library(readxl)
library(ggplot2)
library(ggpubr)

```

```{r merge_whalength_output, echo = F}
# trip ----

trips <- c('2022_07', '2022_10', '2023_01', '2023_05')

merge_trips<-lapply(trips, function(x){
  
  #x<-'2022_07'
  trip<-x
  print(x)

# file paths ----

path <- paste0("./Images/", trip)
date_list <- grep(list.files(path), pattern='.xlsx', invert=TRUE, value=TRUE)

# path to 'screen_grab' folder in each day
sg_path <- lapply(date_list, function(x) paste0(path,"/",x,"/screen grabs"))
# path to 'measure_files' folder in each day
mf_path <- lapply(date_list, function(x) paste0(path,"/",x,"/measure_files"))

# whalelength output merge ----

# list all whalength output csvs
whalength_list <- as.list(unlist(lapply(sg_path, function(x) list.files(x, ".CSV", ignore.case = TRUE, full.names = T))))
#read all whalength csvs without headers
whalength_files <-
  lapply(whalength_list, function(x)
    read.csv(paste0(x), skip = 1)[, 1:50])
#merge all whalength csvs in one dataframe
whalength_merge <- do.call(rbind, whalength_files)
head(whalength_merge)

# IDs and decipher if total length was an actual measurement  ---- 

# list file where I manually add length output to signal that total length in whalength output is real, blank values mean total length is not a real value, but done as hack to measure bh to df insertion
ID_alt<-as.list(unlist(lapply(mf_path, function(x) list.files(x, glob2rx("altperimage*.xlsx"), full.names = T))))

#read all, remove ones with altitude issues, merge
ID_alt_read<-lapply(ID_alt, function(x) readxl::read_excel(x, range = cellranger::cell_cols("A:AI"), na = "NA")%>%filter(issue == "N"))
head(ID_alt_read)

ID_alt_merge <- do.call(rbind, ID_alt_read)

# names of both merged outputs ----

names(whalength_merge)
names(ID_alt_merge)

# merge measurements with ID data ----

length_ID_merge<-whalength_merge%>%
  filter(is.na(Label..for.mult.whales.per.image.))%>%
  left_join(ID_alt_merge, by = c("Filename" = "vlc_filename"))

# add in data where more than one dolphin was measured in one photo ----

if(nrow(whalength_merge%>%
   filter(!is.na(Label..for.mult.whales.per.image.))) > 0){

length_ID_merge<-whalength_merge%>%
  filter(!is.na(Label..for.mult.whales.per.image.))%>%
  left_join(ID_alt_merge, by = c("Filename" = "vlc_filename", "Label..for.mult.whales.per.image." = "ID"))%>%
  bind_rows(length_ID_merge)

}

#has to go at end to work with both single and multi dolphins per image
#removes TL measurement when it is not real
length_ID_merge<-length_ID_merge%>%
  arrange(Filename)%>%
  mutate(trip = trip)%>%
  dplyr::select(trip, ID, everything(), -Whale.ID)%>%
  mutate(Total.Length..m. = case_when(
    is.na(actual_length) ~ NA_real_,
    TRUE ~ Total.Length..m.
  ))

# 0 is NA
length_ID_merge[length_ID_merge == 0] <- NA

length_ID_merge

})

length_ID_merge<-do.call(rbind, merge_trips)


```
```{r ggplot_ID, echo = F}

# check to compare entered TL vs recorded TL in whalength csvs ----
length_ID_merge%>%
  mutate(length_diff = actual_length-Total.Length..m.)%>%
  filter(abs(length_diff)>0.00005)%>%
  dplyr::select(trip, ID, Filename, actual_length,Total.Length..m.,length_diff, png_datetime_nz)

# obtaining the mean value for multiple measurements during the same trip
morphs_trip<-length_ID_merge%>%
  ungroup()%>%
  filter(!is.na(ID))%>%
  group_by(ID, trip)%>%
  #do widths from these individuals as well
  dplyr::summarise(mean_TL = mean(Total.Length..m., na.rm=TRUE), SD_TL = sd(Total.Length..m., na.rm=TRUE),
                   mean_10W = mean(Width.at.10..TL, na.rm=TRUE), SD_10W = sd(Width.at.10..TL, na.rm=TRUE),
                   mean_20W = mean(Width.at.20..TL, na.rm=TRUE), SD_20W = sd(Width.at.20..TL, na.rm=TRUE),
                   mean_30W = mean(Width.at.30..TL, na.rm=TRUE), SD_30W = sd(Width.at.30..TL, na.rm=TRUE),
                   mean_40W = mean(Width.at.40..TL, na.rm=TRUE), SD_40W = sd(Width.at.40..TL, na.rm=TRUE),
                   mean_50W = mean(Width.at.50..TL, na.rm=TRUE), SD_50W = sd(Width.at.50..TL, na.rm=TRUE),
                   mean_60W = mean(Width.at.60..TL, na.rm=TRUE), SD_60W = sd(Width.at.60..TL, na.rm=TRUE),
                   mean_70W = mean(Width.at.70..TL, na.rm=TRUE), SD_70W = sd(Width.at.70..TL, na.rm=TRUE),
                   mean_80W = mean(Width.at.80..TL, na.rm=TRUE), SD_80W = sd(Width.at.80..TL, na.rm=TRUE),
                   mean_90W = mean(Width.at.90..TL, na.rm=TRUE), SD_90W = sd(Width.at.90..TL, na.rm=TRUE),
                   mean_RBH = mean(Rostrum.BH, na.rm=TRUE), SD_RBH = sd(Rostrum.BH, na.rm=TRUE),
                   mean_BHDF = mean(BH.DF.insertion, na.rm=TRUE), SD_BHDF = sd(BH.DF.insertion, na.rm=TRUE),
                   mean_FW = mean(Fluke.width, na.rm = TRUE), SD_FW = sd(Fluke.width, na.rm = TRUE))

#distribution of tl mean vs sd
ggplot(morphs_trip)+
  geom_point(aes(x = mean_TL, y = SD_TL))

#individuals where sd of total length is greater than 10 cm
morphs_trip%>%filter(SD_TL > 0.12)
morphs_trip%>%filter(SD_RBH > 0.1 | SD_BHDF > 0.1)
length_ID_merge%>%filter(ID == 'BIVALVE')
```
# rostrum to bh and bh to df for anyone who had it done
bh_tl<-bh_trip%>%
  left_join(tl_trip, by = "ID")%>%
  filter(!is.na(mean_TL))%>%
  mutate(trip = trip)

#mean
write.csv(bh_tl, paste0(path,"./bh_tl_",trip,".csv"), row.names = F)

#measurement per photo
write.csv(length_ID_merge, paste0(path,"./length_ID_merge_",trip,".csv"), row.names = F)

#tl_bhdf_regression(bh_tl)





```
```{r}

  ggplot(length_ID_merge)+
    geom_point(aes(x = Total.Length..m., y = BH.DF.insertion, color = trip))+
    ggforce::facet_wrap_paginate(~ID, ncol = 3, page = 10)

# # Remove measurements where ID is unknown
# length_ID_merge<-length_ID_merge%>%
#   filter(!is.na(ID))%>%
# 
# 
# 

# 

```
